\documentclass[12pt]{article}
\usepackage[english]{babel}
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\title{MATH-UA 120 Section 8}
\author{Ishan Pranav}
\date{September 16, 2023}
\DeclareMathOperator{\len}{len}
\begin{document}
\maketitle
\section*{List}
A \textit{list} or \textit{tuple} is an ordered sequence of objects. Elements in a list might be repeated.
\section*{Length}
The number of elements in a list $l$ is called its \textit{length} and is denoted $\len(l)$.
\section*{Ordered pair}
A list of length two is called an \textit{ordered pair}.
\section*{Empty list}
A list of length zero is called the \textit{empty list} and is denoted $(\,)$.
\section*{Equal}
Two lists are \textit{equal} provided they have the same length, and elements in the corresponding positions on the two lists are equal. Lists $(a,b,c)$ and $(x,y,z)$ are equal if and only if $a=x,b=y$.
\section*{Concatenation}
Suppose $A$ and $B$ are lists. Their \textit{concatenation} is the new list formed by listing first the elements of $A$ and then the elements $B$. The concatenation of the lists $(1,2,1)$ and $(1,3,5)$ is the list $(1,2,1,1,3,5)$.
\section*{Multiplication principle}
Consider two-element lists for which there are $n$ choices for the first element, and for each choice of the first element, there are $m$ choices for the second element. Then the number of such lists is $nm$. 
\section*{Falling factorial}
The quantity
\[(x)_n=\prod_{k=1}^n{(x-k+1)}=\prod_{k=0}^{n-1}{(x-k)}.\]
is called a \textit{falling factorial}, \textit{descending factorial}, \textit{falling sequential product}, or \textit{lower factorial}. The quantity $(r)_n$ is equal to the number of $r$-permutations from a set of $n$ items, called a \textit{partial permutation} and denoted ${}_nP_r=\frac{n!}{(n-r)!}$.
\section*{List-counting theorem}
The number of lists of length $k$ whose elements are chosen from a pool of $n$ possible elements is $n^k$ if repetitions are permitted and $(n)_k$ if repetitions are forbidden.
\section{All possible two-letter words A, E, I, O, and U}
The possible words are AA, AE, AI, AO, AU, EA, EE, EI, EO, EU, OA, OE, OI, OO, OU, UA, UE, UI, UO, and UU. Of these, 16 have no repeated letters.
\section{The number of different airport codes}
\[3^{26}.\]
\section{The number of length-$k$ bit strings}
\[2^k.\]
\section{Various controls}
\[4\times 3\times 2\times 4=96.\]
\section{Two playlists}
\[\left((500)_{20}\right)^2.\]
\section{3-element lists}
If the first and last entries must be the same: $n^2$.

If the first and last entries must be different: $n^2(n-1)$.
\section{Friends or family}
\begin{enumerate}
\item\[30!\times 2^{30}.\]
\item\[2^{30}.\]
\end{enumerate}
\section{Three different rings}
\end{document}