\documentclass[12pt]{article}
\usepackage[english]{babel}
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\renewcommand{\labelenumi}{\alph{enumi}.}
\title{MATH-UA 120 Section 4}
\author{Ishan Pranav}
\date{September 10, 2023}
\begin{document}
\maketitle
\section*{Pythagorean theorem}
If $a$ and $b$ are the lengths of the legs of a right triangle and $c$ is the length of the hypotenuse, then
\[a^2+b^2=c^2.\]
\section*{Conjecture}
New theorems are the creations of mathematicians that begin as \textit{conjectures}: statements about mathematics whose truth we have yet to establish. Conjectures are educated guesses.
\section{Rewrite}
\begin{enumerate}
    \item The product of an odd integer and an even integer is even. Let $x$ and $y$ be integers. If $x$ is odd and $y$ is even, then $xy$ is even.
    \item The square of an odd integer is odd. If an integer $x$ is odd, then $x^2$ is odd.
    \item The square of a prime number is not prime. If an integer $x$ is prime, then $x^2$ is not prime.
    \item The product of two negative integers is negative. This, of course, is false. Let $x$ and $y$ be integers. If $x$ and $y$ are negative, then $xy$ is negative.
    \item The diagonals of a rhombus are perpendicular. If $\overline{AB}$ and $\overline{CD}$ are the diagonals of a rhombus, then $\overline{AB}\perp\overline{CD}$.
    \item Congruent triangles have the same area. If $\triangle ABC\cong\triangle DEF$, then the area of $\triangle ABC$ is equal to the area of $\triangle DEF$.
    \item The sum of three consecutive integers is divisible by three. If integers $a$, $b$, and $c$ are consecutive (that is, $b=a+1$ and $c=b+1$), then $3|(a+b+c)$.
\end{enumerate}
\section{Indicate}
\begin{enumerate}
    \item If a polygon $PQRS$ is a square, then $PQRS$ is a rectangle.
    \item If a polygon $PQRS$ is a rectangle, then $PQRS$ is a parallelogram.
    \item If Joe is a grandfather, then Joe is male.
    \item If Ellen resides in Los Angeles, then Ellen resides in California.
    \item If this year is a leap year, then this year is divisible by 4.
    \item None.
    \item If $x>0$, then $x^2>0$.
    \item If $X<0$, then $x^3<0$. If $x^3<0$, then $x<0$. Thus, $x<0$ if and only if $x^3<0$.
    \item If $xy=0$, then $x=0$ or $y=0$. If $x=0$ or $y=0$, then $xy=0$. Thus, $xy=0$ if and only if $x=0$ or $y=0$.
    \item If $x=0$ and $y=0$, then $xy=0$.
    \item If $x=0$ and $y=0$, then $x+y=0$.
\end{enumerate}
\section{A common mistake}
Let condition \textit{A} be ``John lives in New York'' and condition \textit{B} be ``John lives in the United States.'' Then, the statement ``if \textit{A}, then \textit{B}'' is true, but the statement ``if \textit{B}, then \textit{A}'' is false.

Let condition \textit{A} be ``A polygon $P$ is a square'' and condition \textit{B} be ``A polygon $P$ is a rectangle.'' Then, the statement ``if \textit{A}, then \textit{B}'' is true, but the statement ``if \textit{B}, then \textit{A}'' is false.
\section{Consider}
The statement ``if \textit{A}, then \textit{B}'' is true unless \textit{A} is true and \textit{B} is false. An ``or'' statement is true unless both conditions are false. This makes the statement identical to the statement ``If (not \textit{A}) or B.''
\section{Consider}
The statement ``if \textit{A}, then \textit{B}'' is identical to the statement ``if (not \textit{B}), then (not \textit{A})'' because the positive and contrapositive are always identical.
\section{Consider}
The statement ''if \textit{A} iff \textit{B}'' is identical to the statement ``(not \textit{A}) iff (not \textit{B}).''
\section{An equilateral triangle}
An equilateral triangle whose side lengths are $a=b=c=1$ is not a violation of the Pythagorean theorem--even though $a^2+b^2\neq c^2$--because the Pythagorean Theorem only necessarily applies to right triangles. By definition, an equilateral triangle is not a right triangle: All angles of an equilateral triangle measure 30 degrees--none measure 90 degrees.
\section{Nonsense}
The statement ``A \textit{line} is the shortest distance between two points'' is nonsense because the definition is incorrect and there is a category mismatch. A better definition is ``A one-dimensional figure $\overline{AB}$ is called a \textit{line segment} provided it connects two points $A$ and $B$, and its length is the distance between $A$ and $B$.
\section{A rather grotesque claim}
The claim ``If you pick a guinea pig up by its tail, then its eyes will pop out'' is mathematically true. This is a vacuous statement: Guinea pigs do not have tails.
\section{The plural of lemma}
The correct plural of \textit{lemma} is \textit{lemmata} (Latin third-declension noun, Greek origin).
\section{Conjectures}
\begin{enumerate}
\item Conjecture: The sums of consecutive odd numbers starting with 1, are perfect squares.
\item Conjecture: The sums of consecutive perfect cubes, starting with 1, are perfect squares.
\end{enumerate}
\end{document}
